---
title: "p8105_hw3_zz2780"
author: "Lydia Zhang"
date: "2022-10-11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .9,
  out.width = "90%"
)
library(patchwork)
```

Problem 1

```{r}
library(p8105.datasets)
data("instacart")
```

The dataset instacart has `r ncol(instacart)` variables and `r nrow(instacart)` observations. Key variables describe different products that are being ordered for each order placed on instacart, including the product name and id, when it is being ordered, the aisle and department it belongs to, and the times it is reordered. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products, from `r instacart%>% select(department) %>% distinct %>% count` departments, found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.


```{r}
instacart %>%
  count(aisle, name="num", sort=TRUE)
```

There are 134 aisles. Fresh vegetables aisle has the most items ordered from.

```{r}
instacart %>%
  count(aisle, name="num")%>%
  filter(num>10000)%>%
  ggplot(aes(x=reorder(aisle, -num), y=num))+
  geom_col()+
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(title="Number of items per Aisle",x="Aisle", y="Number of items purchased")
```

The graph shows number of items ordered from each aisle exluding those with less than 10000 items ordered, arranging in a descending order.

```{r}
instacart%>%
  filter(aisle%in%c("baking ingredients", "dog food care", "packaged vegetables fruits"))%>%
  group_by(aisle)%>%
  count(product_name)%>%
  arrange(desc(n)) %>%
  slice(1:3)
```

Among baking ingredients, light brown sugar, pure baking soda, and cane sugar extract are the three most popular items being ordered 499, 397, and 336 times respectively.

Among dog food care, Snack Sticks Chicken & Rice Recipe Dog Treats, organix chicken & brown rice recipe, Small Dog Biscuits are the three most popular items being ordered 30, 28, and 26 times respectively.

Among packaged vegetable fruits, organic baby spinach, organinc raspberries, and organic blueberries are the three most popular items being ordered 9784, 5546, and 4966 times respectively.

```{r}
instacart%>%
    filter(product_name%in%c("Pink Lady Apples", "Coffee Ice Cream"))%>%
    group_by(product_name, order_dow)%>%
  summarize(mean_hour=mean(order_hour_of_day))%>%
  knitr::kable(digits = 2)
```

The resulting table shows that for each day of the week, the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered.

Problem 2

```{r}
accel<-read_csv("data/accel_data.csv")%>%
  pivot_longer(
    activity.1:activity.1440,
    names_to="minute",
    names_prefix="activity.",
    values_to="counts"
  )%>%
  mutate(
    type=ifelse(day==c("Sunday",  "Saturday"), "Weekend", "Weekday")
  )%>% 
  relocate(type, .before=minute)
accel$minute<-as.numeric(accel$minute)
accel$day<-factor(accel$day, levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))
```

The data set contains `r ncol(accel)` variables and `r nrow(accel)` observations. Variables include the week, day, type of the day (weekend/weekday), and the minute accelerometer collects the activity counts, and the number of counts.

```{r}
accel%>%
  group_by(week, day)%>%
  summarize(sum_counts=sum(counts))
```

The resulting table shows the daily total activity for each week. There is no apparent trend observed.

```{r}
accel%>%
  ggplot(aes(x=minute,y=counts))+
  geom_line(aes(color=day))+
  geom_smooth()+
  viridis::scale_color_viridis(
    option="plasma",
    name = "Day of the Week", 
    discrete = TRUE
  )+
  labs(title="24-hour Activity Plot for Each Day of the Week", y="Activity Counts", x="Minute of the Day")
```

From the graph, we can observe that earlier in the day, individual would have less activity counts, and the number increases as the day goes. During the middle of the day, individual has more activities counts on Sunday that on any other day. As we move onto later in the day, individual has more activities counts on all days in the week except for Sunday. 

Problem 3
```{r}
library(p8105.datasets)
data("ny_noaa")
```

The data set contains `r ncol(ny_noaa)` variables and `r nrow(ny_noaa)` observations. The data contains id for different locations, date of the record, and the temperature, precipitation, and snowfall of that day. The missing data in tmin and tmax will be an issue because tmin and tmax are character variables. Once they are converted to numeric variables the NA will becomes a problem. And if the data has missing value for the entire month or year, calculating the mean of this values for a specific month/year will be a problem.


```{r}
ny_noaa<-ny_noaa%>%
  janitor::clean_names()%>%
  separate(date, into=c("year","month","day"))%>%
  mutate(tmin=as.numeric(tmin), 
         tmax=as.numeric(tmax),
         tmin=tmin/10, 
         tmax=tmax/10,
         prcp=prcp/10,
         year=as.numeric(year),
         month=tolower(month.abb[as.numeric(month)]))
ny_noaa%>%
  count(snow, sort=TRUE)
```

For snowfall, zero is the most common observed value, because there isn't snowing most of the year. 

```{r}
ny_noaa%>%
  filter(month=="jan" | month=="jul")%>%
  group_by(id, month)%>%
  summarise(mean_tmax=mean(tmax))%>%
  na.omit()%>%
  ggplot(aes(x=id,y=mean_tmax))+
  geom_bar(stat='identity')+
  facet_grid(.~month)+
  labs(title="Average Max Temperature in January and July Across Years", x="Station ID", y="Average Max Temperature")
```

I can observe that the average max temperature in January fluctuates around 0 degree Celsius. Only one station has average max temperature high than 5 dgrees Celsius. The average max temperatures in July are all significantly higher than that in January, with most of which falling within the range of 20 to 30 degrees Celsius. There is no significant outlier. 


```{r}
tmax_tmin=
  ny_noaa%>%
  ggplot(aes(x=tmin, y=tmax))+
  geom_hex()+
  labs(title="TMAX vs. TMIN")

snowfall=
  ny_noaa%>%
  filter(snow>0 & snow <100)%>%
  group_by(year)%>%
  summarize(n=n())%>%
  ggplot(aes(x=year, y=n))+
  geom_bar(stat="identity", color="black", fill="blue")+
  labs(title="Distribution of snowfall between 0mm and 100mm across years", x="Year", y="Counts")

tmax_tmin / snowfall

```





